---
title: 'Modeling the impact of vaccination'
output:
  html_document: default
  html_notebook: default
  fig.asp: 0.618
  fig.width: 6
  fig.align: center
  html_notebook: default
  out.width: 70%
  pdf_document: default
editor_options: 
  chunk_output_type: console
---
# Model checking -- population size
```{r}
rm(list=ls())
library( deSolve )
library( tidyverse )
source( "codes/func.R" )
source( "codes/params.R" )
library( Rcpp )
sourceCpp( "codes/cholera_sirw.cpp" )

tstop <- 200
nday <- 365
dt <- 365
times <- seq( 0, tstop*nday, dt )

params <- c( 0.21, 1, 1, 1 )
compute_R0_water( params )
res <- rk( y=init_val, times=times, func=cholera_sirw, parms=params, method="rk45dp7", hmax=4, hmin=0.2 )
if( dt != 365 ){
  thin_row <- seq( 1 , (tstop*nday/dt+1), by=nday/dt )
  res <- res[ thin_row, ]
}
res <- res[ , -1 ]  # remove the first column indicating time
S <- rowSums( res[ , index_s ] )
I <- rowSums( res[ , index_i ] )
R <- rowSums( res[ , index_r ] )
V <- rowSums( res[ , index_v ] )
B <- res[ , index_b ]

N <- S + I + R + V

df <- data.frame( t=0:tstop, S=S, I=I, R=R, V=V, N=N, B=B )
df_long <- tidyr::gather( df, inf_status, num, -t )
# res$inf_status <- factor( res$inf_status, levels=c("S","I","R") )

d <- filter( df_long, inf_status == c("N","S") )
ggplot( d ) +
  geom_line( aes( x=t, y=num, color=inf_status ) ) +
  labs( x="Time (year)", y = "Number", color = "State")

```

# Model checking -- population distribution by age group
```{r}
ag <- lapply( 1:10, function(x) rowSums( res[ , iag[[x]] ] ) )
df <- data.frame( t=0:tstop, ag_0_1=ag[[1]], ag_1_4=ag[[2]], ag_5_14=ag[[3]] )
d <- tidyr::gather( df, age_grp, num, -t )
# d <- filter( df_long, age_grp == "ag_0_1" )
ggplot( d ) +
  geom_line( aes( x=t, y=num, color=age_grp ) ) +
  labs( x="Time (year)", y = "Number", color = "Age group")
```

# Equilibrium distribution of age groups
```{r}
N <- sapply( iag, function(x) rowSums( res[ , x ] ) ) 
tail( N, 1 )

library(rootSolve)
sstate <- runsteady(y=init_val, func=cholera_sirw, parms=params )
sstate$y
sapply( iag, function(x) sum( sstate$y[ x ] ) )
```

# Equilibrium distribution
```{r}
init_val <- dput(tail(out,1)[2:(4*num_age_grp+1)])
init_val <- c( init_val, rep(0,2*num_age_grp) )
dput( init_val )

library(rootSolve)
ss <- runsteady(y=init_val, func=cholera_sir_cpp, parms=params )
```

# Model check -- annual incidence by age group
```{r}
library( tidyverse )
library( deSolve )
source( "codes/params.R" )
source( "codes/func.R" )
library(Rcpp)
sourceCpp( "codes/cholera_sirw.cpp" )

tstop <- 300
times <- seq( 0, tstop*365, 365 )
params <- c( 0.1, 1, 1, 1 )
compute_R0_water( params )
# Integrate ODEs
res <- rk( y=init_val, times=times, func=cholera_sirw, parms=params, method="rk45dp7", hmax=4, hmin=0.1 )
res <- res[ , -1 ] # remove the first column indicating time
S <- res[ , index_s ]
I <- res[ , index_i ] 
R <- res[ , index_r ] 
V <- res[ , index_v ] 
CI <- res[ , index_ci ] 
CV <- res[ , index_cv ] 
B <- res[ , index_b ]
N <- S + I + R + V
annual_inc <- CI[2:nrow(CI),] - CI[1:(nrow(CI)-1),]
pyo <- (N[1:tstop,] + N[2:(tstop+1),])/2 # beginning of the two consecutive years  
annual_inc_per_person <- annual_inc / pyo
# tail( annual_inc_per_person, 1 )
ir <- tail( annual_inc_per_person, 1 )
pop <- tail( N, 1 )
inc_pyo <- ir*1e5
inc_pyo  
  

inc_model( params=params, fun=cholera_sirw, tstop=300, pyo=1e5 )
  
```

# Maximum likelihood parameter estimation
```{r}
rm( list=ls() )
library( deSolve )
source( "codes/params.R" )
source( "codes/func.R" )
library(Rcpp)
library(optimx)
# sourceCpp( "codes/cholera_sir_cpp.cpp" )
sourceCpp( "codes/cholera_sir_smpl_cpp.cpp" )
# tstop <- 600
lower_bounds <- c( 1e-3, 1e-3, 1e-3, 1e-3)
upper_bounds <- c( 1e3, 1e3, 1e3, 1e3 )
start <- c( 1.3, 10, 1, 0.1 )
tic <- Sys.time()
fit <- optimx( par = start, 
               fn = neg_log_lik,
               lower = lower_bounds, 
               upper = upper_bounds,
               method = "nlminb",
               itnmax = 100,
               control = list(trace=5), 
               data = inc_obs_Jakarta )
elapsed <- Sys.time() - tic

params <- sapply( 1:4, function(x) eval(parse(text=paste0("fit$p",x)))) 
# R0_approx = beta*
#   (chi_1*0.01973551 + chi_2*0.07851448 + chi_3*0.1938802 + chi_4*( 0.1886538 + 0.1818183 + 0.1708433 +  0.1497457 +  0.1105314 +  0.05792751 +  0.01713539)
# x <- seq( 0.14, 0.2, 0.01)
# y<- x
# for( i in seq_along(x)){
#   cat( "i = ",i )
#   y[i]<- neg_log_lik( param=x[i], tstop = 200 )
#   cat( ", par = ", x[i], ", nll = ",y[i],"\n")
# }
# fit_list[[i]] <- fit
# 
# obj <- list( fit_results = fit_list, init_pop = y0, par_start = start_pool, t_stop = t_stop )
# obj_name <- paste0( "data/fit_tstop_", t_stop, ".rds" )
# saveRDS( obj, obj_name )
```


# Markov Chain Monte Carlo parameter estimation 
```{r}
rm( list=ls() )
source( "codes/params.R" )
source( "codes/func.R" )
library(Rcpp)
sourceCpp( "codes/cholera_sir_cpp.cpp" )
sourceCpp( "codes/cholera_sir_smpl_cpp.cpp" )
start <- c( 2.8, 1, 0.1, 1 )
tic <- Sys.time()
samp <- run_MCMC( startvalue = start, iter=10000, scale=c(0.005,0.1,0.1,0.1) )
Sys.time() - tic
samp$acceptance_ratio
hist(samp$theta[,1], main = "theta_1")
acf(samp$theta)
```

# Reverse engineering
```{r}
rm( list=ls() )
library( deSolve )
source( "codes/params.R" )
source( "codes/func.R" )
library( Rcpp )
sourceCpp( "codes/cholera_sir_cpp.cpp" )
sourceCpp( "codes/cholera_sir_smpl_cpp.cpp" )
# Generate data
params_true <- c( 1.3, 1, 3, 0.4 )
# ir <- annual_inc_steady( params=params_true )
ir <- annual_inc_tstop( params=params_true, fun=cholera_sir_cpp, tstop=60 )
inc_mean <- ir$annual_inc_pyo*ir$steady_pop/sum(ir$steady_pop)*1e5 # assume a population of 100,000
inc_sample <- rpois( length(inc_mean), inc_mean ) #sample
# reset global variable b/c the log_posterior function needs a global variables, inc_obs
inc_obs[1:3] <- inc_sample[1:3] # <1 yo, 1-4 yo, 5-14 yo
inc_obs[4] <- sum( inc_sample[4:10] ) # 15-64 yo
inc_obs <- round( inc_obs )
start <- c( 0.8, 1, 0.5, 1 )

# tic <- Sys.time()
# mcmc <- run_MCMC( data=inc_obs, fun=cholera_sir_cpp, startvalue=start, iter=10, scale=c(0.01,0.2,0.2,0.2) )
# Sys.time() - tic
# mcmc$acceptance_ratio
# hist(mcmc$theta[,1], main = "theta_1")
# acf(mcmc$theta)

library( adaptMCMC )
mcmc_adt <- MCMC( p=log_posterior, n=100, init=start, scale=c(0.01,0.2,0.2,0.2), adapt=TRUE, acc.rate=0.234, fun=cholera_sir_cpp, data=inc_obs, tstop=60, pyo=1e5, maxsteps=500 )
# tic <- Sys.time()
# Confirm reverse engineering results
d <- readRDS("out/res.rds")
plot(d$samples[,1],type="l")
dd <- d$samples[1e5:2e5,]
apply( dd, 2, mean )
# > apply( dd, 2, mean )
# [1] 1.2361161 0.9054116 3.1729030 0.465244
```

# Fitting to incidence data
```{r}
# library(readxl)
# d <- read_xlsx( "data/cholera_data.xlsx" )
rm( list=ls() )
country <- "Nepal"
library( deSolve )
source( "codes/params.R" )
source( "codes/func.R" )
library( Rcpp )
sourceCpp( "codes/cholera_sirw.cpp" )

start <- c( 0.2, 1, 1, 1 )

library( adaptMCMC )

tic <- Sys.time()
res <- MCMC( p=log_posterior, n=1e5, init=start, scale=c(0.01,0.1,0.1,0.1), adapt=TRUE, acc.rate=0.234, fun=cholera_sirw, data=inc_obs, tstop=150, pyo=1e5 )

# res <- MCMC( p=log_posterior, n=5e4, init=start, scale=c(0.01,0.1,0.1,0.1), adapt=TRUE, acc.rate=0.234, fun=cholera_sirw, data=inc_obs, tstop=200, pyo=1e5, maxsteps=500, hmin=0.1, hmax=4 )

Sys.time() - tic

plot(res$samples[,1],type="l")

tstamp <- format( Sys.time(), "%Y%m%dT%H%M%S" )
saveRDS( res, paste0( "out/", country, "_fit_", tstamp, ".rds") )


# dd <- d$samples[1e5:2e5,]
# apply( dd, 2, mean )
```

# Vaccine impact evaluation
```{r}
rm( list=ls() )
source( "codes/params.R" ) #
source( "codes/func.R" )
library( Rcpp )
sourceCpp( "codes/cholera_sirw.cpp" )
# params <- c(1.3,5,1.8,0.4)

d <- readRDS( "out/res_20190719T132213.rds" )
samples <- d$samples[sample(45000:50000,20),]
scenario <- expand.grid( init_vacc_efficacy = c( 0.95, 0.82 ),
                         waning_rate_vacc_immune = c( 1/(5*365), 1/(20*365) ),
                         frac_report = c( 0.05, 0.1, 0.5 ),
                         vacc_campaign = TRUE )

res_10y <- data.frame( matrix( NA, nrow=nrow(scenario), ncol=nrow(samples) ) )
res_10y <- cbind( scenario, res_10y  )

for( k in 1:nrow(scenario) ){
  cat( "k =", k, "\n" )
  # vacc_campaign <- scenario$vacc_campaign[ k ]
  # cov_vacc_campaign <- scenario$cov_vacc[ k ]
  # start_vacc_campaign <- 90*365
  # stop_vacc_campaign <- 90*365 + dur_vacc_campaign
  dur_post_vacc <- 10
  for( i in 1:nrow(samples) ){
    cat( "i =", i, "\n" )
    res <- incidence( params=samples[ i, 1:4], fun=cholera_sirw, tstop=(dur_post_vacc+90) )
    res_10y[ k, (i+ncol(scenario)) ] <- ( sum(res$ci[(1+dur_post_vacc),]) - sum(res$ci[1,]) )
  }
}

ait <- annual_inc_tstop( params=params, fun=fun, tstop=60*365, dt=1, integ_method="rk45dp7", maxsteps=500 )
ait$annual_inc_pyo
inc <- incidence( params=params, fun=fun, tstart=59*365, tstop=60*365, dt=0.1, integ_method="rk4", maxsteps=500 )
( inc$ci[2,] - inc$ci[1,] ) / inc$pop[1,]
# cumulative incidence over 10 years
target_pop <- 1e5
vacc_campaign <- FALSE
inc <- incidence( params=params, fun=fun, tstart=61*365, tstop=62*365, dt=0.1, integ_method="rk4", maxsteps=500 )
(inc$ci[ nrow(inc$ci), ] - inc$ci[ 1, ]) / sum(inc$pop[1,]) * target_pop

vacc_campaign <- TRUE
cov_vacc_campaign <- rep( 0.9, num_age_grp )
cov_vacc_campaign[1] <- 0  # vaccination except for the first age group
start_vacc_campaign <- 61*365
dur_vacc_campaign <- 14.0
stop_vacc_campaign <- start_vacc_campaign + dur_vacc_campaign
eff_vacc <- rep( 0, nag )
eff_vacc[1:2] <- 0.3 #Bi et al.(2017) Lancet Infect Dis
eff_vacc[3:nag] <- 0.64 #Bi et al.(2017) Lancet Infect Dis
inc <- incidence( params=params, fun=fun, tstart=61*365, tstop=62*365, dt=0.1, integ_method="rk4", maxsteps=500 )
(inc$ci[ nrow(inc$ci), ] - inc$ci[ 1, ]) / sum(inc$pop[1,]) * target_pop

inc <- incidence( params=params, fun=fun, tstart=61*365, tstop=62*365, dt=365, integ_method="rk45dp7", maxsteps=500 )
(inc$ci[ nrow(inc$ci), ] - inc$ci[ 1, ]) / sum(inc$pop[1,]) * target_pop
```

## Inside the incidence method
```{r}
rm( list=ls() )
# steady-state distribution by steate
library(deSolve)
source("codes/params.R") #
source("codes/func.R")
library(Rcpp)
sourceCpp( "codes/cholera_sir_cpp.cpp" )
params <- c(1.3,5,1.8,0.4)
fun <- cholera_sir_cpp

tstart=10
tstop=30
dt=0.1
integ_method="rk4"
maxsteps=500
vacc_campaign <- TRUE
cov_vacc_campaign <- rep( 0.9, num_age_grp )
cov_vacc_campaign[1] <- 0  # vaccination except for the first age group
rate_vacc_campaign <- - log(1-cov_vacc_campaign*eff_vacc) / dur_vacc_campaign
eff_vacc <- rep( 0, nag )
eff_vacc[1:2] <- 0.3 #Bi et al.(2017) Lancet Infect Dis
eff_vacc[3:nag] <- 0.64 #Bi et al.(2017) Lancet Infect Dis
rate_vacc_campaign
start_vacc_campaign <- 5
stop_vacc_campaign <- 19
times <- seq( 0, tstop, dt )
out <- rk( y=init_val, times=times, func=fun, parms=params, method=integ_method, maxsteps=500 )  # Integrate ODEs
row_tstop <- round(tstop/dt)+1
row_tstart <- round(tstart/dt)+1
thin <- seq( 1 , row_tstop, by=round(365/dt) )
out <- out[ thin, ]
S <- out[ , index_S ]
I <- out[ , index_I ]
R <- out[ , index_R ]
V <- out[ , index_V ]
CI <- out[ , index_CI ]
CV <- out[ , index_CV ]
N <- S + I + R + V
  
```


```{r}
library(coda)
chain<- mcmc(samp$samples)
summary(chain)
plot(chain)
library(BayesianTools)
correlationPlot(data.frame(chain))
combinedchains = mcmc.list(chain, chain2)
plot(combinedchains)
gelman.diag(combinedchains)
gelman.plot(combinedchains)

```


```{r}
library(Rcpp)
sourceCpp( "codes/run_mcmc_cpp.cpp" )
```

```{r}
rm( list=ls() )
source( "codes/params.R" ) # includes Jakarta data
source( "codes/func.R" )
library(Rcpp)
sourceCpp( "codes/cholera_sir_smpl_cpp.cpp" )
library(adaptMCMC)
start <- c( 0.3, 1, 1, 1 )
tic <- Sys.time()
samp <- MCMC( p=log_posterior, n=20000, init=start, scale=c(0.01,0.02,0.02,0.02), adapt=TRUE, acc.rate=0.234 )
Sys.time() - tic
```

```{r}
library(coda)
chain<- mcmc(samp$samples)
summary(chain)
plot(chain)
library(BayesianTools)
correlationPlot(data.frame(chain))
combinedchains = mcmc.list(chain, chain2)
plot(combinedchains)
gelman.diag(combinedchains)
gelman.plot(combinedchains)
```

```{r}

```

#Parallel 
```{r}
library(adaptMCMC)
samp <- MCMC( log_posterior, n=1000, init=start, scale=c(0.005,0.1,0.1,0.1), adapt=FALSE)
samp_adt <- MCMC( p=log_posterior, n=2000, init=start, scale=c(0.01,0.2,0.2,0.2), adapt=TRUE, acc.rate=0.234 )
library(coda)
hist( samp_adt$samples[1000:2000,1] )
hist( samp_adt$samples[1000:2000,2] )
hist( samp_adt$samples[1000:2000,3] )
hist( samp_adt$samples[1000:2000,4] )
str(samp)
summary(samp_adt$samples)
plot( samp_adt$samples[,2] )

## ----------------------
## plot density and samples

x1 <- seq(-15, 15, length=80)
x2 <- seq(-15, 15, length=80)
x3 <- seq(-15, 15, length=80)
x4 <- seq(-15, 15, length=80)

d.banana <- matrix(apply(expand.grid(x1, x2), 1,  p.log), nrow=80)

par(mfrow=c(1,2))
image(x1, x2, exp(d.banana), col=cm.colors(60), asp=1, main="no adaption")
contour(x1, x2, exp(d.banana), add=TRUE, col=gray(0.6))
lines(samp.1$samples, type='b', pch=3)

image(x1, x2, exp(d.banana), col=cm.colors(60), asp=1, main="with adaption")
contour(x1, x2, exp(d.banana), add=TRUE, col=gray(0.6))
lines(samp.2$samples, type='b', pch=3)
```


# Steady-state solution

```{r}
rm( list=ls() )
library( deSolve )
source( "codes/params.R" )
source( "codes/func.R" )
library(Rcpp)
sourceCpp( "codes/cholera_sir_cpp.cpp" )
library(rootSolve)
params=c(0.26,1,3,1)

frac_report = 1
rel_susc <- rep(1,nag)
rel_susc[1:3] <- params[2:4]
tstop <- 600
nday <- 365
dt <- 365
times <- seq( 0, tstop*365, dt )

t0 = Sys.time()
y = rk( y=init_val, times=times, func=cholera_sir_cpp, parms=params, method="rk45dp7", hmin=0.2, hmax=4 )
Sys.time() - t0
thin_row <- seq( 1 , (tstop*nday/dt+1), by=nday/dt )
out <- y[ thin_row, ] # make sure that the output appears by year
S <- out[ , index_S ]
I <- out[ , index_I ]
R <- out[ , index_R ]
V <- out[ , index_V ]
CI <- out[ , index_CI ]
CV <- out[ , index_CV ]
N <- S + I + R + V
annual_inc <- CI[2:nrow(CI), ] - CI[1:(nrow(CI)-1),]
pyo <- (N[1:tstop,] + N[2:(tstop+1),])/2 # beginning of the two consecutive years  
annual_inc_pyo <- annual_inc / pyo
tail( annual_inc_pyo, 1 )  
# 
# y2 = runsteady( y=init_val, func=cholera_sir_cpp, parms=params, stol=1e-6, hmin=0.1, hmax=5 )
# out <- y2$y
# frac_report=1
# S <- out[ index_S ]
# I <- out[ index_I ]
# R <- out[ index_R ]
# V <- out[ index_V ]
# N <- S + I + R + V
# foi <- params[1]*sum(I)/sum(N)
# rel_susc <- rep( 1, nag )
# newI <- frac_report*foi*rel_susc*S;
# newI*365/N
#   
sourceCpp( "codes/cholera_sir_smpl_cpp.cpp" )
t0 = Sys.time()
y3 = runsteady( y=init_val[1:(4*nag)], func=cholera_sir_smpl_cpp, parms=params, hmin=0.2, hmax=4 )
Sys.time() - t0
out <- y3$y

S <- out[ index_S-1 ]
I <- out[ index_I-1 ]
R <- out[ index_R-1 ]
V <- out[ index_V-1 ]
N <- S + I + R + V
foi <- params[1]*sum(I)/sum(N)
newI <- frac_report*foi*rel_susc*S;
newI*365/N

```

# Steady-state comparison
```{r}
rm(list=ls())
library( deSolve )
source( "codes/params.R" )
source( "codes/func.R" )
library(Rcpp)
sourceCpp( "codes/cholera_sir_cpp.cpp" )
library(rootSolve)
sourceCpp( "codes/cholera_sir_smpl_cpp.cpp" )

# d <- matrix(nrow=50,ncol=4)
nr <- 20
params_pool <- matrix( nrow=nr, ncol=4 )
params_pool[,1] <- rnorm( nr, 0.3, 0.2 )
params_pool[,2] <- rnorm( nr, 1, 0.5 )
params_pool[,3] <- rnorm( nr, 1, 0.5 )
params_pool[,4] <- rnorm( nr, 1, 0.5 )
rel_susc <- matrix(1, nrow=nr, ncol=10 )

# hist(R0)
get_all_positive_row <- function( mat ){
  rows_selected <- sapply( 1:nrow(mat), function(x) all(sign(mat[x,])>0)  )
  mat[rows_selected,]
  
}
params <- get_all_positive_row( params_pool )
# compute 
rel_susc[,1:3] <- params[,2:4]
R0 <- rep(NA,nrow(rel_susc))
for( i in 1:nrow(rel_susc)){
  R0[i] <- params_pool[i,1]/gamma*sum(rel_susc[i,]*age_dist)
}
params <- params[(R0>1&R0<10),]
paramsR0 <- cbind( params, R0[(R0>1&R0<10)] )

d <- matrix(NA, nrow=nrow(params), ncol=4)
# hist(R0)
inc_index <- 5
for( i in 1:nrow(d) ){
  cat( "i=", i, "\n" )
  par <- as.double(params[i, ])
  frac_report = 1
  rel_susc <- rep(1,nag)
  rel_susc[1:3] <- par[2:4]
  tstop <- 600
  nday <- 365
  dt <- 365
  times <- seq( 0, tstop*365, dt )
  
  t0 = Sys.time()
  y = rk( y=init_val, times=times, func=cholera_sir_cpp, parms=par, method="rk45dp7", hmin=0.2, hmax=4 )
  elapsed <- Sys.time() - t0
  thin_row <- seq( 1 , (tstop*nday/dt+1), by=nday/dt )
  out <- y[ thin_row, ] # make sure that the output appears by year
  S <- out[ , index_S ]
  I <- out[ , index_I ]
  R <- out[ , index_R ]
  V <- out[ , index_V ]
  CI <- out[ , index_CI ]
  CV <- out[ , index_CV ]
  N <- S + I + R + V
  annual_inc <- CI[2:nrow(CI), ] - CI[1:(nrow(CI)-1),]
  pyo <- (N[1:tstop,] + N[2:(tstop+1),])/2 # beginning of the two consecutive years  
  annual_inc_pyo <- annual_inc / pyo
  tail( annual_inc_pyo, 1 )  
  d[i,1] <- tail( annual_inc_pyo, 1 )[inc_index]  
  d[i,2] <- elapsed 
  
  t0 = Sys.time()
  y3 = runsteady( y=init_val[1:(4*nag)], func=cholera_sir_smpl_cpp, parms=par, hmin=0.2, hmax=5 )
  elapsed <- Sys.time() - t0
  out <- y3$y
  
  S <- out[ index_S-1 ]
  I <- out[ index_I-1 ]
  R <- out[ index_R-1 ]
  V <- out[ index_V-1 ]
  N <- S + I + R + V
  foi <- par[1]*sum(I)/sum(N)
  newI <- frac_report*foi*rel_susc*S;
  annual_newI <- newI*365/N
  d[i,3] <- annual_newI[inc_index]  
  d[i,4] <- elapsed 
}
# dd <- data.frame(id=1:nrow(d), d=d)
# df <- tbl_df(dd)
# df_long <- gather( df[,c(2,4)], key, value, - id )
# ggplot( df )+
#   geom_line( aes(meth))
```



```{r}
# out <- rk( y=yini, times=t_steps, func=derivs, parms=pars, method="rk45ck" )
# > class(out)
# [1] "deSolve" "matrix" 
class(out) <- "matrix" # without this, dplyr functions get error
out <- tbl_df( out )
colnames( out ) <- c( "t", "S", "I", "R" )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

out <- filter( out, inf_status == "I" )

ggplot( out ) + 
    geom_line( aes( x=t, y=num, color=inf_status ) ) +
    geom_line( aes( x=t, y=num ), color="red" ) +
    scale_colour_manual( values=c(S="blue",I="red",R="green") ) +
    scale_y_continuous( limits=c(0,520) ) +
    scale_x_continuous( limits=c(0,30) ) +
    labs( x="Time (year)", y="Number of infected people" )
    labs( x="Time (year)", y="Number of infected people", color="Infection status" )
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "sir_i_bir_seas_10-24yr.png" )
```

### 1. Solving the logistic equation that describes population growth
```{r}
library( tidyverse )
K <- 10
r <- 1  #  dy/dt = r*y*(1-y/K)
y0 <- 20

derivs <- function( t, y, pars ){
    list( r * y * (1-y/K) )
}

library(deSolve)
times <- seq( 0, 20, 0.2 )

out <- ode( y=y0, times=times, func = derivs, parms = NULL )

class( out ) <- "matrix"
out <- tbl_df( out )
colnames( out ) <- c( "t", "y" )

ggplot( out ) + 
    geom_line( aes(x=t,y=y ) ) +
    labs( x="Time (day)", y="Number of people" ) + 
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("r == ", r), parse=T, x=17, y=9, size=4, colour="red" ) +
    annotate( "text", label=paste0("y0 == ", y0), parse=T, x=19, y=9, size=4, colour="red" ) 
```

### 2. Solving the Lorenz model
```{r}
library( tidyverse )

derivs <- function( t, y, pars ){
    X <- y[ 1 ]
    Y <- y[ 2 ]
    Z <- y[ 3 ]
    a <- pars[ 1 ]
    b <- pars[ 2 ]
    c <- pars[ 3 ]
    
    dXdt = a * X + Y * Z
    dYdt = b * ( Y - Z )
    dZdt = - X*Y + c*Y - Z;
    
    list( c( dXdt, dYdt, dZdt) )
}

library(deSolve)
times <- seq( 0, 100, 0.01 )
pars <- c( a=-8/3, b=-10, c=28 )
y0 <- c( X=1, Y=1, Z=1 )

out <- ode( y=y0, times=times, func = derivs, parms = pars )

class( out ) <- "matrix"
out <- tbl_df( out )
out <- gather( out, var, val, -time )
ggplot( out ) + 
    geom_line( aes( x=time,y=val, color=var ) ) +
    labs( x="Time (day)", y="Number of people" )

ggplot( out ) + 
    geom_line( aes( x=time,y=val, color=var ) ) +
    labs( x="Time (day)", y="Number of people" )


```


### 2. Solving ODEs for SIR model by manually writing the forward Euler method
```{r}
library( tidyverse )
beta <- 0.2
gamma <- 0.1

dt <- 1
t_end <- 200
t_steps <- seq( 0, t_end, by=dt )
t_num <- length(t_steps)
S <- I <- R <- rep( NA, t_num ) 
S0 <- 999
I0 <- 1
R0 <- 0

S[ 1 ] <- S0
I[ 1 ] <- I0
R[ 1 ] <- R0

for( i in 1 : (t_num-1) ){
    N <- S[ i ] + I[ i ] + R[ i ] 
    dS <- - beta * S[ i ] * I [i ] / N
    dI <- beta * S[ i ] * I[ i ] / N - gamma * I[ i ]
    dR <- gamma * I[ i ]
    
    S[ i+1 ] <- S[ i ] + dS * dt ## forward Euler method: y1 = y0 + h*f(x0,y0), y' = f(x0,y0)
    I[ i+1 ] <- I[ i ] + dI * dt 
    R[ i+1 ] <- R[ i ] + dR * dt 
    
}

out <- data_frame( t=t_steps, S=S, I=I, R=R )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

ggplot( out ) + 
    geom_line( aes(x=t,y=num, color=inf_status ) ) +
    labs( x="Time (day)", y="Number of people", color="Infection status" ) + 
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "dt.pdf" )
```

### 3. Solve ODEs using the deSolve package
```{r}
library( tidyverse )
library( deSolve )

derivs <- function( t, y, pars ){
    S <- y[ 1 ]
    I <- y[ 2 ]
    R <- y[ 3 ]
    N <- S + I + R
    beta <- pars[1]
    gamma <- pars[2]
    
    dS <- - beta * S * I / N
    dI <- beta * S * I / N - gamma * I
    dR <- gamma * I
    
    return( list( c( dS, dI, dR ) ) )
    
}
beta <- 0.2
gamma <- 0.1
pars <- c( beta, gamma )
t_end <- 200
dt <- 0.1
t_steps <- seq( 0, t_end, dt )
S0 <- 999
I0 <- 1
R0 <- 0
yini <- c( S0, I0, R0 )
out <- ode( y=yini, times=t_steps, func=derivs, parms=pars )
# out <- rk( y=yini, times=t_steps, func=derivs, parms=pars, method="rk45ck" )
# > class(out)
# [1] "deSolve" "matrix" 
class(out) <- "matrix" # without this, dplyr functions get error
out <- tbl_df( out )
colnames( out ) <- c( "t", "S", "I", "R" )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

out <- filter( out, inf_status == "I" )
ggplot( out ) + 
    geom_line( aes( x=t, y=num, color=inf_status ) ) +
    scale_colour_manual( values=c(S="blue",I="red",R="green") ) +
    scale_y_continuous( limits=c(0,1000) ) +
    labs( x="Time (day)", y="Number of people", color="Infection status" ) + 
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "sir_i.png" )
# ggsave( "sir_i.pdf" )
```

### 3. Simulation of a measles epidemic
```{r}
library( tidyverse )
library( deSolve )

derivs <- function( t, y, pars ){
    S <- y[ 1 ]
    I <- y[ 2 ]
    R <- y[ 3 ]
    N <- S + I + R
    beta0 <- pars[1]
    beta1 <- pars[2]
    c <- pars[3]
    gamma <- pars[4]
    nu <- pars[5]
    beta <- beta0 * (1 + beta1*sin(2*pi*(t + c)) )
    v <- pars[6]
    if( t < 15 ) { v <- 0 } # vaccination takes place only after year 15
    dS <- - beta * S * I / N + nu*N*(1-vacc) - nu*S
    dI <- beta * S * I / N - gamma * I - nu*I
    dR <- gamma * I - nu*R + nu*N*vacc
    
    return( list( c( dS, dI, dR ) ) )
    
}
# biennial epidemics
beta <- 0.5*365
gamma <- (1/15)*365
R <- beta/gamma
nu <- 1/15
beta0 <- beta
beta1 <- 8.8
c <- 0.25
vacc <- 0.3
# beta <- 1.0*365
# gamma <- 0.1*365
# nu <- 0/10
# beta0 <- beta
# beta1 <- 0
# c <- -0.75
pars <- c( beta0, beta1, c, gamma, nu, vacc  )
t_end <- 50
dt <- 0.002
t_steps <- seq( 0, t_end, dt )
S0 <- 9990
I0 <- 10
R0 <- 0
yini <- c( S0, I0, R0 )
out <- ode( y=yini, times=t_steps, func=derivs, parms=pars )
# out <- rk( y=yini, times=t_steps, func=derivs, parms=pars, method="rk45ck" )
# > class(out)
# [1] "deSolve" "matrix" 
class(out) <- "matrix" # without this, dplyr functions get error
out <- tbl_df( out )
colnames( out ) <- c( "t", "S", "I", "R" )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

out <- filter( out, inf_status == "I" )

ggplot( out ) + 
    # geom_line( aes( x=t, y=num, color=inf_status ) ) +
    geom_line( aes( x=t, y=num ), color="red" ) +
    # scale_colour_manual( values=c(S="blue",I="red",R="green") ) +
    scale_y_continuous( limits=c(0,520) ) +
    scale_x_continuous( limits=c(0,30) ) +
    labs( x="Time (year)", y="Number of infected people" )
    # labs( x="Time (year)", y="Number of infected people", color="Infection status" ) 
    # theme( legend.position=c(0.85,0.45) ) +
    # annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    # annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "sir_i_bir_seas_10-24yr.png" )
```



### 3-1. Modeling through generational basis
```{r}
library( tidyverse )
beta <- 0.2
gamma <- 0.1
R0 <- 2
dt <- 1
median_inf_period <- 6.738122
t_end <- 200
t_steps <- seq( 0, t_end, by=median_inf_period )
t_num <- length(t_steps)
S <- I <- R <- rep( NA, t_num ) 
S_init <- 999
I_init <- 1
R_init <- 0

S[ 1 ] <- S_init
I[ 1 ] <- I_init
R[ 1 ] <- R_init

for( i in 1 : (t_num-1) ){
    N <- S[ i ] + I[ i ] + R[ i ] 
    
    new_inf <- S[ i ] * ( 1 - exp( - R0 * I [ i ] / N ) )

    S[ i+1 ] <- S[ i ] - new_inf## forward Euler method: y1 = y0 + h*f(x0,y0), y' = f(x0,y0)
    I[ i+1 ] <- new_inf
    R[ i+1 ] <- R[ i ] + I[ i ] 
    
}

out <- data_frame( t=t_steps, S=S, I=I, R=R )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

ggplot( out ) + 
    geom_line( aes(x=t,y=num, color=inf_status ) ) +
    labs( x="Time (day)", y="Number of people", color="Infection status" ) + 
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "gen.png" )

```
### 3. Cholera transmission
```{r}
cholera_model <- function( t, y, params ){
    source( "data/params.R" )
    S <- y[ 1 : num_age_grp ]
    I <- y[ (1+num_age_grp) : (2*num_age_grp) ]
    R <- y[ (1+2*num_age_grp) : (3*num_age_grp) ]
    V <- y[ (1+3*num_age_grp) : (4*num_age_grp) ]
    # 
    dSdt <- rep( 0, num_age_grp )
    dIdt <- rep( 0, num_age_grp )
    dRdt <- rep( 0, num_age_grp )
    dVdt <- rep( 0, num_age_grp )
    
    
    beta <- params[ 1 ]
    gamma <- params[ 2 ]
    
    # Derived variables
    N <- sum( S ) + sum( I ) + sum( R ) + sum( V )
    mu <- death_rate
    births <- birth_rate / sum(sample_Y0)
    
    foi <- beta * sum(I) / N
    dCVdt <- rep( 0, num_age_grp )
    dCIdt <- rep( 0, num_age_grp )
    
	  for( i in  1:num_age_grp ) {
	    if( i == 1 ){
	      dSdt[i] = + births - (foi*rel_susc[i] + mu[i] + vacc_rate_campaign[i] + ag[i])*S[i]
  	    dIdt[i] = + foi*rel_susc[i]*S[i] - (mu[i] + ag[i] + gamma)*I[i]
  	    dRdt[i] = + (1 - case_fatality[i])*gamma*I[i] - (mu[i] + ag[i] + vacc_rate_campaign[i])*R[i]
  	    dVdt[i] = + vacc_rate_campaign[i]*(S[i] + R[i]) - (mu[i] + ag[i])*V[i]		
	    } else {
	      dSdt[i] = + ag[i-1]*S[i-1] - (foi*rel_susc[i] + mu[i] + vacc_rate_campaign[i] + ag[i])*S[i]
	      dIdt[i] = + ag[i-1]*I[i-1] + foi*rel_susc[i]*S[i] - (mu[i] + ag[i] + gamma)*I[i]
	      dRdt[i] = + ag[i-1]*R[i-1] + (1 - case_fatality[i])*gamma*I[i] - (mu[i] + ag[i] + vacc_rate_campaign[i])*R[i]
	      dVdt[i] = + ag[i-1]*V[i-1] + vacc_rate_campaign[i]*(S[i] + R[i]) - (mu[i] + ag[i])*V[i]
	    }
	    dCVdt[i] <- vacc_rate_campaign[i]*(S[i] + R[i]) # cumulative number of vaccine doses by age group
	    dCIdt[i] = foi*rel_susc[i]*S[i] # cumulative incidence by age group
	  }
    
    dydt <- c( dSdt, dIdt, dRdt, dVdt, dCIdt, dCVdt )
    list( dydt )
}
```
<!-- ### 4. Calling an agent-based SIR model written in Java using the rJava package -->
<!-- ```{r} -->
<!-- library( rJava ) -->
<!-- library( tidyverse ) -->

<!-- .jinit( force.init=TRUE ) # starts a Java virtual machine -->
<!-- path_parallel_colt_lib <- "C:/Program Files/Java/lib/ParallelColt/parallelcolt-0.9.4.jar" -->
<!-- path_sir_agent_model <- "C:/Users/jonghoon.kim/Documents/Workspace/Eclipse/SIR_Agent_Based/bin" -->
<!-- .jaddClassPath( path_sir_agent_model ) -->
<!-- .jaddClassPath( path_parallel_colt_lib ) -->
<!-- pars <- .jnew( "Parameters" ) -->
<!-- model <- .jnew( "Model" ) -->

<!-- .jcall( pars, "V", "setNumInitInfected", 9L ) -->
<!-- .jcall( pars, "V", "setTransmissionRate", 0.3 ) -->

<!-- beta <- .jcall( pars, "D", "getTransmissionRate" ) -->
<!-- gamma <- .jcall( pars, "D", "getRecoveryRate" ) -->

<!-- arr <- .jcall( model, "[[D", "runModel", pars ) -->
<!-- out <- sapply( arr, .jevalArray ) -->
<!-- out <- t( out ) -->

<!-- out <- tbl_df( out ) -->
<!-- colnames( out ) <- c( "t", "S", "I", "R" ) -->
<!-- out <- gather( out, inf_status, num, -t ) -->
<!-- out$inf_status <- factor( out$inf_status, levels=c("S","I","R") ) -->

<!-- ggplot( out ) +  -->
<!--     geom_line( aes( x=t, y=num, color=inf_status ) ) + -->
<!--     labs( x="Time (day)", y="Number of people", color=NULL ) +  -->
<!--     theme( legend.position=c(0.90,0.45) ) + -->
<!--     annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) + -->
<!--     annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=120, y=1000, size=4, colour="red" ) -->

<!-- # ggsave( "sir_agent_based.png" ) -->

<!-- ``` -->



#### Q#1: Estimate the $\beta$ and $\gamma$ using the SIR model
(Taken from p.17 of "An Introduction to Mathematical Epidemiology" by Maia Martcheva)

In January and February 1978, an epidemic of influenza occurred in a boarding
school in the North of England. The boarding school housed a total of 763 boys, all
of whom were at risk during the epidemic. The spring term began on January 10.
The boys returned from their Christmas vacation spent at many different locations in
the world. A boy returning from Hong Kong exhibited elevated temperature during
the period 15-18 January. On January 22, three boys were sick. Table 1 gives the
number of boys ill on the $n$th day beginning January 22 ($n=1$).

Table 1. Daily number of influenza-infected boys

| Day | No. infected<sup>a</sup> | Day  | No. infected<sup>a</sup>  |
| --- |------------:| -- | ---------: |
| 3   | 25             | 9  |  192    |
| 4   | 75             | 10 |  126    |
| 5   | 227            | 11 |  71     |
| 6   | 296            | 12 |  28     |
| 7   | 258            | 13 |  11     |
| 8   | 236            | 14 |  7      |
<sup>a</sup>Data taken from "Influenza in a Boarding School," British Medical Journal, 4 March 1978



