---
title: 'Modeling the impact of vaccination'
output:
  html_document: default
  html_notebook: default
  fig.asp: 0.618
  fig.width: 6
  fig.align: center
  html_notebook: default
  out.width: 70%
  pdf_document: default
editor_options: 
  chunk_output_type: console
---
# Model checking--total population size (R version)
```{r}
library( tidyverse )
library( deSolve )
source( "data/params.R" )
source( "cholera_sirw.R" ) # cholera model
tstop <- 55
times <- seq( 0, tstop*365, 10 )
params <- c( beta=0.4, chi_1=1, chi_2=1, chi_3=1, chi_4=1 )
# Integrate ODEs
out <- ode( y=sample_init_val, times=times, func=cholera_sirw_r, parms=params )

S <- rowSums( out[ , 2 : (num_age_grp+1) ] )
I <- rowSums( out[ , (2+num_age_grp) : (2*num_age_grp+1) ] )
R <- rowSums( out[ , (2+2*num_age_grp) : (3*num_age_grp+1) ] )
V <- rowSums( out[ , (2+3*num_age_grp) : (4*num_age_grp+1) ] )
N <- S + I + R + V
df <- data.frame( t=(out[,1]/365),S=S, I=I, R=R, V=V, N=N )
df_long <- tidyr::gather( df, inf_status, num, -t )
# out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

# d <- filter( df_long, inf_status == "I" )
ggplot( df_long ) + 
    geom_line( aes( x=t, y=num, color=inf_status ) ) 

```

# Model check -- population size (Rcpp)
```{r}
library( tidyverse )
library( deSolve )
source( "data/params.R" )
library(Rcpp)
sourceCpp( "cholera_sirw_cpp.cpp" )

tstop <- 60
times <- seq( 0, tstop*365, 10 )
params <- c( beta=0.25, chi_1=1, chi_2=1, chi_3=1, chi_4=1 )
# Integrate ODEs
# out <- ode( y=sample_init_val, times=times, func=cholera_sirw_cpp, parms=params 
out <- rk( y=init_val, times=times, func=cholera_sirw_cpp, parms=params, method="rk45dp7" )

S <- rowSums( out[ , 2 : (num_age_grp+1) ] )
I <- rowSums( out[ , (2+num_age_grp) : (2*num_age_grp+1) ] )
R <- rowSums( out[ , (2+2*num_age_grp) : (3*num_age_grp+1) ] )
V <- rowSums( out[ , (2+3*num_age_grp) : (4*num_age_grp+1) ] )
N <- S + I + R + V
df <- data.frame( t=(out[,1]/365),S=S, I=I, R=R, V=V, N=N )
df_long <- tidyr::gather( df, inf_status, num, -t )
# out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

# d <- filter( df_long, inf_status == "I" )
ggplot( df_long ) +
    geom_line( aes( x=t, y=num, color=inf_status ) )

```

# Equilibrium distribution to the initial condition
```{r}
init_val <- dput(tail(out,1)[2:(4*num_age_grp+1)])
init_val <- c( init_val, rep(0,2*num_age_grp) )
dput( init_val )
```

# Model check -- annual incidence by age group
```{r}
library( tidyverse )
library( deSolve )
source( "data/params.R" )
library(Rcpp)
sourceCpp( "cholera_sirw_cpp.cpp" )

tstop <- 60
times <- seq( 0, tstop*365, 365 )
params <- c( beta=0.25, chi_1=1, chi_2=1, chi_3=1, chi_4=1 )
# Integrate ODEs
out <- rk( y=init_val, times=times, func=cholera_sirw_cpp, parms=params, method="rk45dp7" )
nag <- num_age_grp
S <- out[ , 2 : (nag+1) ]
I <- out[ , (2+nag) : (2*nag+1) ]
R <- out[ , (2+2*nag) : (3*nag+1) ]
V <- out[ , (2+3*nag) : (4*nag+1) ]
CI <- out[ ,(2+4*nag) : (5*nag+1) ]
CV <- out[ ,(2+5*nag) : (6*nag+1) ]
N <- S + I + R + V
annual_CI <- CI[2:nrow(CI),] - CI[1:(nrow(CI)-1),]
pyo <- (N[1:tstop,] + N[2:(tstop+1),])/2 # beginning of the two consecutive years  
  
annual_CI_pyo <- annual_CI / pyo
tail(annual_CI_pyo)  

```

# Annual incidence 
```{r}
# calculate the incidence at the final year 
annual_incidence <- function( FUN = NULL, params=NULL, init_val=NULL, tstop=NA ){
    if( !is.function(FUN) ) 
      stop("argument FUN is not a function!")
   # to get yearly output as the unit time in the model is 365 days
    source( "data/params.R" )
    if( is.na(tstop) )
      tstop <- 60
    if( is.null(params) )
      params <- c( beta=0.3, gamma=1/5 )
    if( is.null(init_val) )
      init_val <- sample_init_val / sum(sample_init_val)
      
    times <- seq( 0, tstop*365, 365 )
    # Integrate ODEs
    out <- ode( y=init_val, times=times, func=FUN, parms=params )
    nag <- num_age_grp
    S <- out[ , 2 : (nag+1) ]
    I <- out[ , (2+nag) : (2*nag+1) ]
    R <- out[ , (2+2*nag) : (3*nag+1) ]
    V <- out[ , (2+3*nag) : (4*nag+1) ]
    CI <- out[ ,(2+4*nag) : (5*nag+1) ]
    N <- S + I + R + V
    annual_CI <- CI[2:nrow(CI),] - CI[1:(nrow(CI)-1),]
    pyo <- (N[1:tstop,] + N[2:(tstop+1),])/2 # beginning of the two consecutive years  
    annual_CI_pyo <- annual_CI / pyo
    tail( annual_CI_pyo, 1)  
}  
```


```{r}
# neg_log_lik
# function to calculate sum of negative log likelihood
# input is set to have one compoent, par, to ease the use of optim function
neg_log_lik <- function( params=NULL, data=NULL, init_val=NULL ){
  stopifnot( length(params) == 5, length(data) == 8, !is.null(init_val_java), !is.null(run_java) )

  inc_model <- annual_incidence(FUN=cholera_sirw_cpp, params=params, tstop=tstop  )
  # assume Poisson distribution 
  neg_log_lik <- - sum( dpois( data, lambda=inc_rate_model, log=TRUE ) )
}
```

# parameter estimation
```{r}
rm( list=ls() )

source( "data/params.R" )


t_stop <- 80


start_pool <- seq( 1,50, length.out=5 )
fit_list <- list()
lower_bounds <- rep( 1E-24, 5 )
upper_bounds <- c( 1000, 10, 1, 1000, 1000 )
for( i in 5:5 ) {
# start <- c( 1, 0.1, 0.1, 1, 1 )
start <- c( beta=start_pool[i], chi_1=0.1, chi_2=0.5, chi_3=0.01, chi_4=0.1 )

cat( "i=", i, "\n")
fit <- optimx::optimx( par = start, 
                  fn = neg_log_lik,
                  lower = lower_bounds, 
                  upper = upper_bounds,
                  method = "nlminb",
                  # itnmax = 100,
                  control = list(trace=3),
                  data = inc_rate_obs, 
                  init_val = init_val )

fit_list[[i]] <- fit
}
obj <- list( fit_results = fit_list, init_pop = y0, par_start = start_pool, t_stop = t_stop )
obj_name <- paste0( "data/fit_tstop_", t_stop, ".rds" )
saveRDS( obj, obj_name )
```


```{r}
# out <- rk( y=yini, times=t_steps, func=derivs, parms=pars, method="rk45ck" )
# > class(out)
# [1] "deSolve" "matrix" 
class(out) <- "matrix" # without this, dplyr functions get error
out <- tbl_df( out )
colnames( out ) <- c( "t", "S", "I", "R" )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

out <- filter( out, inf_status == "I" )

ggplot( out ) + 
    geom_line( aes( x=t, y=num, color=inf_status ) ) +
    geom_line( aes( x=t, y=num ), color="red" ) +
    scale_colour_manual( values=c(S="blue",I="red",R="green") ) +
    scale_y_continuous( limits=c(0,520) ) +
    scale_x_continuous( limits=c(0,30) ) +
    labs( x="Time (year)", y="Number of infected people" )
    labs( x="Time (year)", y="Number of infected people", color="Infection status" )
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "sir_i_bir_seas_10-24yr.png" )
```

### 1. Solving the logistic equation that describes population growth
```{r}
library( tidyverse )
K <- 10
r <- 1  #  dy/dt = r*y*(1-y/K)
y0 <- 20

derivs <- function( t, y, pars ){
    list( r * y * (1-y/K) )
}

library(deSolve)
times <- seq( 0, 20, 0.2 )

out <- ode( y=y0, times=times, func = derivs, parms = NULL )

class( out ) <- "matrix"
out <- tbl_df( out )
colnames( out ) <- c( "t", "y" )

ggplot( out ) + 
    geom_line( aes(x=t,y=y ) ) +
    labs( x="Time (day)", y="Number of people" ) + 
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("r == ", r), parse=T, x=17, y=9, size=4, colour="red" ) +
    annotate( "text", label=paste0("y0 == ", y0), parse=T, x=19, y=9, size=4, colour="red" ) 
```

### 2. Solving the Lorenz model
```{r}
library( tidyverse )

derivs <- function( t, y, pars ){
    X <- y[ 1 ]
    Y <- y[ 2 ]
    Z <- y[ 3 ]
    a <- pars[ 1 ]
    b <- pars[ 2 ]
    c <- pars[ 3 ]
    
    dXdt = a * X + Y * Z
    dYdt = b * ( Y - Z )
    dZdt = - X*Y + c*Y - Z;
    
    list( c( dXdt, dYdt, dZdt) )
}

library(deSolve)
times <- seq( 0, 100, 0.01 )
pars <- c( a=-8/3, b=-10, c=28 )
y0 <- c( X=1, Y=1, Z=1 )

out <- ode( y=y0, times=times, func = derivs, parms = pars )

class( out ) <- "matrix"
out <- tbl_df( out )
out <- gather( out, var, val, -time )
ggplot( out ) + 
    geom_line( aes( x=time,y=val, color=var ) ) +
    labs( x="Time (day)", y="Number of people" )

ggplot( out ) + 
    geom_line( aes( x=time,y=val, color=var ) ) +
    labs( x="Time (day)", y="Number of people" )


```


### 2. Solving ODEs for SIR model by manually writing the forward Euler method
```{r}
library( tidyverse )
beta <- 0.2
gamma <- 0.1

dt <- 1
t_end <- 200
t_steps <- seq( 0, t_end, by=dt )
t_num <- length(t_steps)
S <- I <- R <- rep( NA, t_num ) 
S0 <- 999
I0 <- 1
R0 <- 0

S[ 1 ] <- S0
I[ 1 ] <- I0
R[ 1 ] <- R0

for( i in 1 : (t_num-1) ){
    N <- S[ i ] + I[ i ] + R[ i ] 
    dS <- - beta * S[ i ] * I [i ] / N
    dI <- beta * S[ i ] * I[ i ] / N - gamma * I[ i ]
    dR <- gamma * I[ i ]
    
    S[ i+1 ] <- S[ i ] + dS * dt ## forward Euler method: y1 = y0 + h*f(x0,y0), y' = f(x0,y0)
    I[ i+1 ] <- I[ i ] + dI * dt 
    R[ i+1 ] <- R[ i ] + dR * dt 
    
}

out <- data_frame( t=t_steps, S=S, I=I, R=R )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

ggplot( out ) + 
    geom_line( aes(x=t,y=num, color=inf_status ) ) +
    labs( x="Time (day)", y="Number of people", color="Infection status" ) + 
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "dt.pdf" )
```

### 3. Solve ODEs using the deSolve package
```{r}
library( tidyverse )
library( deSolve )

derivs <- function( t, y, pars ){
    S <- y[ 1 ]
    I <- y[ 2 ]
    R <- y[ 3 ]
    N <- S + I + R
    beta <- pars[1]
    gamma <- pars[2]
    
    dS <- - beta * S * I / N
    dI <- beta * S * I / N - gamma * I
    dR <- gamma * I
    
    return( list( c( dS, dI, dR ) ) )
    
}
beta <- 0.2
gamma <- 0.1
pars <- c( beta, gamma )
t_end <- 200
dt <- 0.1
t_steps <- seq( 0, t_end, dt )
S0 <- 999
I0 <- 1
R0 <- 0
yini <- c( S0, I0, R0 )
out <- ode( y=yini, times=t_steps, func=derivs, parms=pars )
# out <- rk( y=yini, times=t_steps, func=derivs, parms=pars, method="rk45ck" )
# > class(out)
# [1] "deSolve" "matrix" 
class(out) <- "matrix" # without this, dplyr functions get error
out <- tbl_df( out )
colnames( out ) <- c( "t", "S", "I", "R" )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

out <- filter( out, inf_status == "I" )
ggplot( out ) + 
    geom_line( aes( x=t, y=num, color=inf_status ) ) +
    scale_colour_manual( values=c(S="blue",I="red",R="green") ) +
    scale_y_continuous( limits=c(0,1000) ) +
    labs( x="Time (day)", y="Number of people", color="Infection status" ) + 
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "sir_i.png" )
# ggsave( "sir_i.pdf" )
```

### 3. Simulation of a measles epidemic
```{r}
library( tidyverse )
library( deSolve )

derivs <- function( t, y, pars ){
    S <- y[ 1 ]
    I <- y[ 2 ]
    R <- y[ 3 ]
    N <- S + I + R
    beta0 <- pars[1]
    beta1 <- pars[2]
    c <- pars[3]
    gamma <- pars[4]
    nu <- pars[5]
    beta <- beta0 * (1 + beta1*sin(2*pi*(t + c)) )
    v <- pars[6]
    if( t < 15 ) { v <- 0 } # vaccination takes place only after year 15
    dS <- - beta * S * I / N + nu*N*(1-vacc) - nu*S
    dI <- beta * S * I / N - gamma * I - nu*I
    dR <- gamma * I - nu*R + nu*N*vacc
    
    return( list( c( dS, dI, dR ) ) )
    
}
# biennial epidemics
beta <- 0.5*365
gamma <- (1/15)*365
R <- beta/gamma
nu <- 1/15
beta0 <- beta
beta1 <- 8.8
c <- 0.25
vacc <- 0.3
# beta <- 1.0*365
# gamma <- 0.1*365
# nu <- 0/10
# beta0 <- beta
# beta1 <- 0
# c <- -0.75
pars <- c( beta0, beta1, c, gamma, nu, vacc  )
t_end <- 50
dt <- 0.002
t_steps <- seq( 0, t_end, dt )
S0 <- 9990
I0 <- 10
R0 <- 0
yini <- c( S0, I0, R0 )
out <- ode( y=yini, times=t_steps, func=derivs, parms=pars )
# out <- rk( y=yini, times=t_steps, func=derivs, parms=pars, method="rk45ck" )
# > class(out)
# [1] "deSolve" "matrix" 
class(out) <- "matrix" # without this, dplyr functions get error
out <- tbl_df( out )
colnames( out ) <- c( "t", "S", "I", "R" )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

out <- filter( out, inf_status == "I" )

ggplot( out ) + 
    # geom_line( aes( x=t, y=num, color=inf_status ) ) +
    geom_line( aes( x=t, y=num ), color="red" ) +
    # scale_colour_manual( values=c(S="blue",I="red",R="green") ) +
    scale_y_continuous( limits=c(0,520) ) +
    scale_x_continuous( limits=c(0,30) ) +
    labs( x="Time (year)", y="Number of infected people" )
    # labs( x="Time (year)", y="Number of infected people", color="Infection status" ) 
    # theme( legend.position=c(0.85,0.45) ) +
    # annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    # annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "sir_i_bir_seas_10-24yr.png" )
```



### 3-1. Modeling through generational basis
```{r}
library( tidyverse )
beta <- 0.2
gamma <- 0.1
R0 <- 2
dt <- 1
median_inf_period <- 6.738122
t_end <- 200
t_steps <- seq( 0, t_end, by=median_inf_period )
t_num <- length(t_steps)
S <- I <- R <- rep( NA, t_num ) 
S_init <- 999
I_init <- 1
R_init <- 0

S[ 1 ] <- S_init
I[ 1 ] <- I_init
R[ 1 ] <- R_init

for( i in 1 : (t_num-1) ){
    N <- S[ i ] + I[ i ] + R[ i ] 
    
    new_inf <- S[ i ] * ( 1 - exp( - R0 * I [ i ] / N ) )

    S[ i+1 ] <- S[ i ] - new_inf## forward Euler method: y1 = y0 + h*f(x0,y0), y' = f(x0,y0)
    I[ i+1 ] <- new_inf
    R[ i+1 ] <- R[ i ] + I[ i ] 
    
}

out <- data_frame( t=t_steps, S=S, I=I, R=R )
out <- gather( out, inf_status, num, -t )
out$inf_status <- factor( out$inf_status, levels=c("S","I","R") )

ggplot( out ) + 
    geom_line( aes(x=t,y=num, color=inf_status ) ) +
    labs( x="Time (day)", y="Number of people", color="Infection status" ) + 
    theme( legend.position=c(0.85,0.45) ) +
    annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) +
    annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=122, y=1000, size=4, colour="red" )

# ggsave( "gen.png" )

```
### 3. Cholera transmission
```{r}
cholera_model <- function( t, y, params ){
    source( "data/params.R" )
    S <- y[ 1 : num_age_grp ]
    I <- y[ (1+num_age_grp) : (2*num_age_grp) ]
    R <- y[ (1+2*num_age_grp) : (3*num_age_grp) ]
    V <- y[ (1+3*num_age_grp) : (4*num_age_grp) ]
    # 
    dSdt <- rep( 0, num_age_grp )
    dIdt <- rep( 0, num_age_grp )
    dRdt <- rep( 0, num_age_grp )
    dVdt <- rep( 0, num_age_grp )
    
    
    beta <- params[ 1 ]
    gamma <- params[ 2 ]
    
    # Derived variables
    N <- sum( S ) + sum( I ) + sum( R ) + sum( V )
    mu <- death_rate
    births <- birth_rate / sum(sample_Y0)
    
    foi <- beta * sum(I) / N
    dCVdt <- rep( 0, num_age_grp )
    dCIdt <- rep( 0, num_age_grp )
    
	  for( i in  1:num_age_grp ) {
	    if( i == 1 ){
	      dSdt[i] = + births - (foi*rel_susc[i] + mu[i] + vacc_rate_campaign[i] + ag[i])*S[i]
  	    dIdt[i] = + foi*rel_susc[i]*S[i] - (mu[i] + ag[i] + gamma)*I[i]
  	    dRdt[i] = + (1 - case_fatality[i])*gamma*I[i] - (mu[i] + ag[i] + vacc_rate_campaign[i])*R[i]
  	    dVdt[i] = + vacc_rate_campaign[i]*(S[i] + R[i]) - (mu[i] + ag[i])*V[i]		
	    } else {
	      dSdt[i] = + ag[i-1]*S[i-1] - (foi*rel_susc[i] + mu[i] + vacc_rate_campaign[i] + ag[i])*S[i]
	      dIdt[i] = + ag[i-1]*I[i-1] + foi*rel_susc[i]*S[i] - (mu[i] + ag[i] + gamma)*I[i]
	      dRdt[i] = + ag[i-1]*R[i-1] + (1 - case_fatality[i])*gamma*I[i] - (mu[i] + ag[i] + vacc_rate_campaign[i])*R[i]
	      dVdt[i] = + ag[i-1]*V[i-1] + vacc_rate_campaign[i]*(S[i] + R[i]) - (mu[i] + ag[i])*V[i]
	    }
	    dCVdt[i] <- vacc_rate_campaign[i]*(S[i] + R[i]) # cumulative number of vaccine doses by age group
	    dCIdt[i] = foi*rel_susc[i]*S[i] # cumulative incidence by age group
	  }
    
    dydt <- c( dSdt, dIdt, dRdt, dVdt, dCIdt, dCVdt )
    list( dydt )
}
```
<!-- ### 4. Calling an agent-based SIR model written in Java using the rJava package -->
<!-- ```{r} -->
<!-- library( rJava ) -->
<!-- library( tidyverse ) -->

<!-- .jinit( force.init=TRUE ) # starts a Java virtual machine -->
<!-- path_parallel_colt_lib <- "C:/Program Files/Java/lib/ParallelColt/parallelcolt-0.9.4.jar" -->
<!-- path_sir_agent_model <- "C:/Users/jonghoon.kim/Documents/Workspace/Eclipse/SIR_Agent_Based/bin" -->
<!-- .jaddClassPath( path_sir_agent_model ) -->
<!-- .jaddClassPath( path_parallel_colt_lib ) -->
<!-- pars <- .jnew( "Parameters" ) -->
<!-- model <- .jnew( "Model" ) -->

<!-- .jcall( pars, "V", "setNumInitInfected", 9L ) -->
<!-- .jcall( pars, "V", "setTransmissionRate", 0.3 ) -->

<!-- beta <- .jcall( pars, "D", "getTransmissionRate" ) -->
<!-- gamma <- .jcall( pars, "D", "getRecoveryRate" ) -->

<!-- arr <- .jcall( model, "[[D", "runModel", pars ) -->
<!-- out <- sapply( arr, .jevalArray ) -->
<!-- out <- t( out ) -->

<!-- out <- tbl_df( out ) -->
<!-- colnames( out ) <- c( "t", "S", "I", "R" ) -->
<!-- out <- gather( out, inf_status, num, -t ) -->
<!-- out$inf_status <- factor( out$inf_status, levels=c("S","I","R") ) -->

<!-- ggplot( out ) +  -->
<!--     geom_line( aes( x=t, y=num, color=inf_status ) ) + -->
<!--     labs( x="Time (day)", y="Number of people", color=NULL ) +  -->
<!--     theme( legend.position=c(0.90,0.45) ) + -->
<!--     annotate( "text", label=paste0("beta == ", beta), parse=T, x=100, y=1000, size=4, colour="red" ) + -->
<!--     annotate( "text", label=paste0("gamma == ", gamma), parse=T, x=120, y=1000, size=4, colour="red" ) -->

<!-- # ggsave( "sir_agent_based.png" ) -->

<!-- ``` -->



#### Q#1: Estimate the $\beta$ and $\gamma$ using the SIR model
(Taken from p.17 of "An Introduction to Mathematical Epidemiology" by Maia Martcheva)

In January and February 1978, an epidemic of influenza occurred in a boarding
school in the North of England. The boarding school housed a total of 763 boys, all
of whom were at risk during the epidemic. The spring term began on January 10.
The boys returned from their Christmas vacation spent at many different locations in
the world. A boy returning from Hong Kong exhibited elevated temperature during
the period 15-18 January. On January 22, three boys were sick. Table 1 gives the
number of boys ill on the $n$th day beginning January 22 ($n=1$).

Table 1. Daily number of influenza-infected boys

| Day | No. infected<sup>a</sup> | Day  | No. infected<sup>a</sup>  |
| --- |------------:| -- | ---------: |
| 3   | 25             | 9  |  192    |
| 4   | 75             | 10 |  126    |
| 5   | 227            | 11 |  71     |
| 6   | 296            | 12 |  28     |
| 7   | 258            | 13 |  11     |
| 8   | 236            | 14 |  7      |
<sup>a</sup>Data taken from "Influenza in a Boarding School," British Medical Journal, 4 March 1978



